---
title: "Tools & Best Practices for the Integration of Geospatial Data "
output: html_document
---
```{r,eval=F,echo=F}
rm(list=ls(all=T))
```

## Overview
This notebook was developed to accompany the tutorial of a short course offered at the 2017 Annual Meeting of the American Political Science Association. The instructors for the course are Karsten Donnay (University of Konstanz), Eric Dunford (University of Maryland), Andrew Linke (University of Utah), Erin McGrath (University of Maryland), David Backer (University of Maryland), and David Cunningham (University of Maryland). This short course focuses on newly developed software tools designed by the instructors, which enable more effective work with multiple datasets that have geospatial properties, which are increasingly employed in research conducted throughout the social sciences. The aims of the course are to familiarize participants with the use of these tools and associated best practices. At the end of the course, participants should understand why and how they could use these tools to support relevant research that requires integrating datasets with particular geospatial properties.

The first part of the notebook walks through the functionality, applications and best practices of the `geomerge` package, which was just released. This package has been designed primarily to facilitate addressing challenges related to the integration of datasets with different geospatial properties. The package is illustrated using example data for Nigeria 2011. The illustration covers integration of `Polygon`, `Raster` and `Point` data, including how to generate spatial panel data.

The first part of the notebook walks through the functionality, applications and best practices of the `meltt` package, which was released earlier this year. This package has been designed to facilitate the integration of event data from multiple sources with differing properties. The package is illustrated by drawing on conflict event data from four prominent event datasets covering conflict observed in Nigeria during 2011.

The tutorial is designed to be hands-on, with participants working through the illustrative examples, accessing and processing datasets using the commands available in the `geomerge` and `meltt` packages. Doing so requires, at a minimum, an installation of the R programming software. Some knowledge of R is useful, though not mandatory. During this short course and tutorial, participants should learn about the utility, logic, and functionality of the two packages even without any significant expertise in R.


# Part 1: geomerge - Geospatial Data Integration

## Why we developed geomerge

The use case: In practice, research involving spatial data typically entails drawing on multiple sources that provide information on distinct variables, each with a particular geographical resolution. Conducting analysis requires integrating these variables from the separate datasets into a common data frame, with a geographical resolution that is appropriately comparable across all the variables.

The main challenges: Separate datasets can have very different spatial data formats. For example, information on population or elevation is most often available as `Raster` data. Information on a country's administrative subdivisions is typically provided as `Polygon` data. The locations of conflict events or incidents of crime are usually coded as `Point` data. In essence, these data formats correspond to different units of observation. Different units implies a spatial mismatch. When spatial data are mismatched, they may not be usable for particular types of analysis (unless purposely considering variables at different units of observation). Separate datasets may also treat the same variable as being of different types (e.g., numeric vs. categorical). 

The technical challenges: A whole range of packages in `R` provide excellent functionality for dealing with these data integration problems, without a single, simple framework that combines all this functionality. In addition, integrating different kinds of spatial data requires making assumptions and providing specifications for how to proceed with the integration.

## What geomerge Does

The `geomerge` package provides this framework. The package allows for the automatic, flexible, transparent, reproducible integration of the most common types of spatial data. The integration can produce variables with the same spatial resolution, or merely establish the spatial correspondence of variables with different resolutions. In doing so, the package implements a number of established best practices that ensure robust results for many standard cases, while allowing for customization through optional parameters.

## Using geomerge in Research

`geomerge` supports empirical research using spatial data in several important ways. First, the package streamlines the process of integrating data from multiple sources. Second, the package offers the flexibility of enabling users to generate variants of the same data. Each of these variants can reflect different assumptions about how to perform the integration, including in reference to the choice of spatial unit, as well as the choice of assignment, zonal function or point aggregation rules. Third, the variants can be used to test the robustness of analyses to assumptions about data integration. Fourth, the package contributes to transparency and simplifies replication by providing clear, standardized interfaces that document the assumptions users made when integrating data. The data and code used in performing any integration can be supplied to accompany similar code used in performing analysis.


## Installation
The package can be installed through the **CRAN repository**.

```{r,eval=F}
#install.packages("geomerge")
```

We recommend to install the latest development version of the package from Github for the purposes of this tutorial. To download this version, you may first need to install the `devtools` package. 

```{r,eval=F}
# install.packages("devtools")
devtools::install_github("css-konstanz/geomerge")
```

```{r}
library(geomerge)
```

Before we get started, please set your work directory to the directory into which you unpacked the tutorial files (including the "data" directory).

```{r,eval=F}
setwd("YOUR DIRECTORY") 
```

## Example Data: Nigeria
In this tutorial, we use a number of different data layers for Nigeria 2011 that constitute the example data distributed with the `geomerge` package. The data can be easily loaded using

```{r}
data(geomerge)
```

The example datasets cover all three main spatial data types discussed above:

- `ACLED` (`Point` data): Conflict events for Nigeria in 2011 as recorded by the Armed Conflict Location & Event Data project, available from http://www.acleddata.com/data. This dataset contains geocoded, timestamped information on individual conflict events.

- `AidData` (`Point` data, including locations geocoded to administrative divisions, but assigned coordinates of centroids): Activities of development aid projects in Nigeria with start dates in 2011 as recorded by AidData, available at http://aiddata.org. This dataset contains geocoded, timestamped information on individual aid projects.

Note: Both `Point` datasets are time-stamped, which means that they can be used for dynamic (i.e., spanning a spatial panel) as well as static (i.e., cross-sectional) integration.

- `geoEPR` (`Polygons` data): All politically relevant ethnic groups for Nigeria in 2011, as recorded in the EPR-Core 2014 dataset, available at https://icr.ethz.ch/data/epr/geoepr/. This dataset assigns every politically relevant ethnic group one of six settlement patterns and provides polygons describing their location.

- `gpw` (`Raster` data): Population at a gridded resolution of about 4km for Nigeria in 2010, as compiled by CIESIN, available at http://sedac.ciesin.columbia.edu/data/collection/gpw-v4. This dataset provides population estimates at several grid resolutions.

- `states` (`Polygons` data): Second-order administrative divisions (ADM2s) for Nigeria, known as Local Government Areas (equivalent of US states). The dataset is available at http://www.arcgis.com/home/item.html?id=0e58995046b74254911c1dc0eb756fa4. 
This dataset is used in the illustration for the target SpatialPolygonsDataFrame to which spatial data are merged. The polygons in `states` have been simplified to reduce the size of the `SpatialPolygonsDataFrame` and enable fast execution of the examples provided.


To familiarize yourself with these datasets, we recommend to take a closer look at them. To see a handful of sample values for each:

```{r}
library(raster)
# Quick overview plot
plot(states)
# Show top rows of dataset
head(states@data)
```
```{r}
# Quick overview plot
plot(states)
plot(ACLED, new=TRUE,add=TRUE)
# Show top rows of dataset
head(ACLED@data)
```
```{r}
# Quick overview plot
plot(states)
plot(AidData, new=TRUE,add=TRUE)
# Show top rows of dataset
head(AidData@data)
```
```{r}
# Quick overview plot
plot(geoEPR)
# Show top rows of dataset
head(geoEPR@data)
```

## Using geomerge

The main functionality of the `geomerge` package is provided by a single function with the same name. The output of the function is an object of class "geomerge", which is a list with three slots: (1) `data` contains the spatial data resulting from integration, (2) `inputData` stores the input dataset, and (3) `parameters` logs all parameters with which `geomerge` was executed.

Running `geomerge` has two basic requirements.

The first requirement is input data, comprised of any number of objects of type `SpatialPolygonsDataFrame`, `SpatialPointsDataFrame` and `RasterLayer`. The `RasterLayer` will always by definition be single-valued. Therefore, `geomerge` requires the user to select one specific variable in each of the `SpatialPolygonsDataFrame` and `SpatialPointsDataFrame` objects prior to integration. `SpatialPointsDataFrame` may also contain a second column named `timestamp`, which can be used for dynamic integration.

The rationale is that the package uses the name of the input data to label the corresponding variables in the integrated data. This approach establishes a clear, unique link between the input and integrated data. If a user wishes to work with several variables from the same dataset, simply enter these variables as separate arguments (with unique names). We generally advise users to rely on meaningful names when labeling input data.

The second requirement, called `target`, specifies the spatial structure to which variables from all input objects are merged. The example in the `geomerge` package requires this `target` to be of class `SpatialPolygonsDataFrame`. In practice, the spatial structure can have any shape (e.g., polygons of administrative units, raster cells, etc.).

Note: The package provides a useful helper function called `integrateGrid`, which generates a grid of user-specified cell size for the spatial extent defined by a spatial `R` object.

`geomerge` assumes that all inputs of type `SpatialPolygonsDataFrame` and `RasterLayer` are static and contemporary. If the polygons or raster are changing, we advise users to rerun `geomerge` for each interval in which data are static and contemporary. The package allows for dynamic integration of all inputs that are a `SpatialPointsDataFrame`. For example, one can automatically generate the counts of events that occur within a specific unit of `target` within a specific time period.

`geomerge` has a number of other optional arguments, which we will explore further below. These optional arguments enable specific kinds of integration (i.e., dynamic vs. static) and/or allow the user to change assumptions about zonal functions, assignment rules, etc. from the default values.

Note: The `print`, `summary` and `plot` functions are overloaded for objects of class "geomerge", meaning that these functions return specific outputs for objects of class "geomerge".

## Static Integration of Polygon and Raster Data

The simplest case is that of merging static layers. Consider, for example, the case that geo-spatial information about the settlement areas of ethnic groups ought to be merged with the administrative units of a country to determine which group is the dominant faction in each area. In the following examples, we therefore assume that the `target` of integration is the `states` `SpatialPolygonsDataFrame`.

Let's begin by integrating one `Polygon` dataset with `states`.
```{r}
output = geomerge(geoEPR,target=states)
summary (output)
names(output$data)
```
Notice that the function returns a number of messages documenting the progress of the integration task. When merging more complex data, the function may run for some time and monitoring progress can therefore be relevant. If no printed progress updates are required, simply use the optional argument `silent = TRUE`.

```{r}
output = geomerge(geoEPR,target=states,silent=TRUE)
summary (output)
```

Here, the default settings of `geomerge` make implicit assumptions regarding the assignment of the values in `geoEPR` to the `target` of `states` `SpatialPolygonsDataFrame`. The default assignment rule uses maximum area overlap (`assignment = "max(area)"`). This rule implies that a value is assigned to any spatial unit of `target` that corresponds to the unit in `geoEPR` with the largest spatial overlap.

As an alternative, `geomerge` supports assignment based on minimal area overlap (`assignment = "min(area)"`).

Assignment can also be done by maximum population (`assignment = "max(pop)"`) or minimum population (`assignment = "min(pop)"`), which operate similar to the area .

In addition, `geomerge` permits assignment weighted by area (`assignment = "weighted(area)"`) or population (`assignment = "weighted(pop)"`). The former assigns the value that is the area-weighted average across all units intersecting with the spatial unit in `target`. The latter is analogous, but assigns the value based on the population represented by that area.

Naturally, all the options relying on population require a population raster input called `population.data`. Here is an example:

```{r}
output = geomerge(geoEPR,target=states,
                  silent=TRUE,assignment="max(pop)",
                  population.data=gpw)
summary (output)
```

Note: Any weighted assignment (whether area- or population-based) is only allowed for numeric data. Within our illustration, therefore, weighted assignment is not possible for the layer `geoEPR`.

The integration of `Raster` data is similarly straightforward.

Note: `geomerge` accepts any optional arguments of the function `extract` in the `raster` package. These arguments can be entered in the exact same syntax as in the original `extract` function and are passed on to any use of the function within the package. For example, in the illustration we use the optional input `na.rm = TRUE` because the `gpw` data has a few missing values that we want to ignore when performing the data integration.

```{r,fig.height=4,fig.width=9}
output = geomerge(gpw,na.rm=TRUE,target=states)
summary (output)
plot(output)
```

As can be seen in the `summary`, the package not only merged the layer `gpw` to `states`, but also generated its value per area of the target polygon and first- and second-order spatial lag values for each. For inputs of type `RasterLayer`, values per area are always also returned. Whether or not spatial lags should be calculated can be controlled by the optional Boolean argument `spat.lag`.

```{r,fig.height=4,fig.width=9}
output = geomerge(gpw,na.rm=TRUE,target=states,spat.lag=FALSE)
summary (output)
plot(output)
```

As in the case of `Polygon` data, the defaults of `geomerge` have built-in implicit assumptions regarding zonal statistics. The default zonal function is summation (`zonal.fun = sum`). The package also supports all zonal statistics consistent with the `extract` function in the `raster` package.

```{r,fig.height=4,fig.width=9}
output = geomerge(gpw,na.rm=TRUE,target=states,
                  spat.lag=FALSE,zonal.fun=min)
summary (output)
plot(output)
```

## Static and Dynamic Integration of Point Data

In `geomerge`, integration of `point` data supports two different heuristics, which the user specifies via `point.agg`. The first heuristic (`point.agg = "cnt"`) counts the occurrence of points in a given unit of `target`. The second heuristic users (`point.agg = "sum"`) sums the values for all points in a given unit. This heuristic is only appropriate for numeric variables.

To illustrate, we use information on the conflict fatalities as recorded in `ACLED` and the financial commitments of development aid projects as recorded in `AidData`. We start by looking at the event counts and the number of projects in each Local Government Area of Nigeria throughout 2011 using `point.agg = "cnt"`. Then we examine the total numbers of conflict fatalities and aid dollar commitments associated with those areas.

```{r,fig.height=12,fig.width=7}
# First select the corresponding columns only
ACLED.fatalities = ACLED[,names(ACLED)=='FATALITIES']
AidData.commitment = AidData[,names(AidData)=='commitme_1']
```
```{r}
# Run geomerge using point.agg = 'cnt
output = geomerge(ACLED.fatalities,AidData.commitment,target=states,point.agg='cnt')
summary(output)
plot(output)
```
```{r}
# Run geomerge using point.agg = 'sum
output = geomerge(ACLED.fatalities,AidData.commitment,target=states,point.agg='sum')
summary(output)
plot(output)
```

Dynamic integration of `point` data follows the same process as before, but separated in a series of temporal units, thereby generating a spatial panel. In `geomerge`, the temporal units are specified through the `time` argument. The package performs static integration if `time = NA`. For dynamic integration, the user must specify `time = c(start_date, end_date, interval_length).` All three inputs must be `strings`, where `interval_length` is defined in multiples of `t_unit`. The default value is `t_unit = "days"`. The package also accepts inputs of "secs", "mins", "hours", "months" or "years".

In the following illustration, we employ the same data as before, but now include the "timestamp" column from both datasets. Information capturing the timing of observations is a prerequisite for dynamic integration. The information does not have to be at any specific level of precision, but does have to concern timing. We iterate through the whole year 2011 in one-month steps. In other words, we generate a county-month spatial panel.

```{r,fig.height=12,fig.width=7}
# First select the corresponding columns only
ACLED.fatalities = ACLED[,names(ACLED)%in%c('timestamp','FATALITIES')]
AidData.commitment = AidData[,names(AidData)%in%c('timestamp','commitme_1')]
```
```{r}
# Run geomerge using point.agg = 'cnt
output = geomerge(ACLED.fatalities,AidData.commitment,
                  target=states,time=c("2011-01-01","2011-12-31","1"),
                  t_unit='months',point.agg='cnt')
summary(output)
plot(output)
```
```{r}
# Run geomerge using point.agg = 'cnt
output = geomerge(ACLED.fatalities,AidData.commitment,
                  target=states,time=c("2011-01-01","2011-12-31","1"),
                  t_unit='months',point.agg='sum')
summary(output)
plot(output)
```

Note: By default, `plot` selects the last time period for purposes of the visualization. If the user wishes to visualize any other period, simply add the optional argument `period` to the function. Also, first- and second-order time-lagged variables are returned by default. The optional Boolean argument `time.lag` controls this feature.

```{r}
output = geomerge(ACLED.fatalities,AidData.commitment,
                  target=states,time=c("2011-01-01","2011-12-31","1"),
                  t_unit='months',point.agg='sum',time.lag=FALSE)
summary(output)
plot(output, period=3)
```

## Generating Grid Target

Thus far, we have only considered integration targets in the form of the Nigeria county polygons `states`. The `generateGrid` function in `geomerge` allows the user to easily generate a matching grid of a chosen resolution. For many econometric applications, this option can be very useful.

```{r,fig.height=4,fig.width=9}
# install.packages("sp")
require(sp)

# Generate grid with 10 km cell size (input in m) in local CRS for Nigeria
states.grid <- generateGrid(states,
                              size= 10000, # meters
                              local.CRS=CRS("+init=epsg:26391"),
                              silent = TRUE)

# Run simple static integration with this grid as target
output = geomerge(ACLED.fatalities,target=states.grid,point.agg='sum')
summary(output)
plot(output)
```

